### 2018-10-08 preamble.tex
reformatted to R Journal style

% Any extra LaTeX you need in the preamble
\usepackage{amsmath}