### 2019-05-17 cover-letter/cover-letter.Rmd
tweak cover letter and reponse letter

---
from: Earo Wang
fromqualifications: B.Comm.(Hons)
fromtitle: PhD candidate
fromWWW: earo.me
fromemail: earo.wang@monash.edu
fromphone: +61 4 4919 1031
fromsignature: signature.png
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Sincerely
spacing: 1.4
output: MonashEBSTemplates::letter
---

We have undertaken major revision on our manuscript titled, "Calendar-based graphics for visualizing people's daily schedules," based on reviewers' comments.

The major changes include two new sections: a subsection "2.3.5 Faceted calendar" to describe a new calendar faceting method `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`, and a subsection "2.4 Reasons to use calendar-based graphics" to explain why calendar displays are useful and when to use them.

A detailed response letter to reviewers has been provided.

We believe that this manuscript has been significantly improved, that is a good fit for the Journal of Computational and Graphical Statistics.