### 2019-10-28 cover-letter/cover-letter.Rmd
added cover letter

---
from: Earo Wang
fromqualifications: B.Comm.(Hons)
fromtitle: PhD candidate
fromWWW: earo.me
fromemail: earo.wang@monash.edu
fromphone: +61 4 4919 1031
fromsignature: signature.png
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Sincerely
spacing: 1.4
output: MonashEBSTemplates::letter
---

We have done minor revision on our manuscript titled, "Calendar-based graphics for visualizing people's daily schedules," based on associate editor and reviewers' comments.

A detailed response letter to reviewers has been provided.

We believe that this manuscript has been improved, that is a good fit for the Journal of Computational and Graphical Statistics.